---
title: "A01-Logistic_on_final_exam_data"
output:
  html_notebook: default
---

Author: Joss Ives, *updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*.

# Setup

## Load libraries
```{r}
suppressWarnings(library(lme4))
require(lme4)
suppressWarnings(library(ggplot2))
require(ggplot2)
source("jossfunc.R")
```

## Data files

### Load data
```{r}
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat.csv")
```

### Add additional calculated values
```{r}
dat.raw$course.grade.frac <- dat.raw$course.grade/100.
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$final.grade.LMH <- as.integer(
  cut(dat.raw$f.tot78, quantile(dat.raw$f.tot78, probs=0:3/3), include.lowest=TRUE)
  )
dat.raw$final.grade.fix <- (dat.raw$f.tot78-2.*dat.raw$QCORRECT)/76.
dat.raw$final.gradeA.fix <- (dat.raw$f.Atot40-2.*dat.raw$QCORRECT)/38.

names(dat.raw)
```


### Set categorical variables
```{r}
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$QNUM      <- factor(dat.raw$QNUM)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)

```

### Make some data subsets
```{r}
# Keep only those data points where TREATMENT was set
dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Look only at the 4 questions that had TREATMENT
dat.trt <- subset(dat.all, QNUM==5 | QNUM==6 | QNUM==9 | QNUM==10)
names(dat.trt)
```



# Initial analyses on 4 EYA questions

### What do the results look like for each question?
```{r}
dodge=position_dodge(width=0.9)

corr.by.question <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","QNUM")
  )

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.question, aes(x=QNUM, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Question number", y = "Fraction correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  theme(axis.text = element_text(face = "bold")) + 
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  guides(fill=guide_legend(title=NULL)) 
  #ggtitle("All")
```


### Fisher's Exact for overall question set
```{r}
## fisher.2vector
# 2 vectors of 0s and 1s are passed
# * The first is control (top row)
# * The second is treatement (bottom row)
# Adding parenthesis around assigning a variable causes the output to be displayed
(result <- fisher.2vector(
  dat.trt$QCORRECT[dat.trt$TREATMENT==0],
  dat.trt$QCORRECT[dat.trt$TREATMENT==1]
)) 
cat("Cohen's d\n ", cohens.d.from.odds.simple(result$estimate[[1]]),"\n")

```

### Simple logistic regression
```{r}
m <- glmer(QCORRECT ~ QNUM + TREATMENT + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```


### A quick summary of odds ratios so far
```{r}
odds.table <- matrix(c(result$estimate[[1]],exp(tab[[5]])),ncol=1,byrow=TRUE)
colnames(odds.table) <- c("Odds Ratio")
rownames(odds.table) <- c("Fisher's Exact Test","Logistic Regression")
as.table(odds.table)
```

# Digging into the logistic regressions a bit more

### QNUM as a random effect
First, let's look at making QNUM a random effect. Not sure if it is correct, but it makes no difference to the results and cleans up the reporting.
```{r}
m.2re <- glmer(QCORRECT ~ (1|QNUM) + TREATMENT + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.2re)$coefficients
```

```{r}
summary(m.2re)
```


### Fitting using all the questions

This makes the AIC and BIC quadruple and doesn't provide a significant change in TREATMENT1. 

```{r}
#m.allQs <- glmer(QCORRECT ~ QNUM + TREATMENT + (1|ID), 
#                data = dat.all, 
#                family = binomial, control=glmerControl(optimizer="bobyqa"))
#summary(m.allQs)
```

```{r}
#m.allQs.2re <- glmer(QCORRECT ~ (1|QNUM) + TREATMENT + (1|ID), 
#                data = dat.all, 
#                family = binomial, control=glmerControl(optimizer="bobyqa"))
#summary(m.allQs.2re)
```

# Looking at other variables with only the treatment questions

## Course grade

Course grade and NCRT both are significant and improve the model slightly, but do not have a significant impact on TREATMENT1
```{r}
m.grade <- glmer(QCORRECT ~ QNUM + TREATMENT + course.grade.frac + NCRT + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.grade)
```

# Time to dig into NCRT

## Graphs

### EYA performance vs NCRT and median split

```{r}
s.ncrt <- summarySE(dat.trt, measurevar="QCORRECT",
                              groupvars=c("NCRT","TREATMENT"), .drop=FALSE)
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
#options(repr.plot.width=5, repr.plot.height=3)

ggplot(s.ncrt, aes(x=NCRT, y=QCORRECT, fill=TREATMENT)) + 
    geom_bar(stat="identity", position=dodge) + 
    geom_errorbar(limits, position=dodge, width=0.5)
```

```{r}
s.crtsplit <- summarySE(dat.trt, measurevar="QCORRECT",
                              groupvars=c("CRT.medsplit","TREATMENT"), .drop=FALSE)
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
#options(repr.plot.width=5, repr.plot.height=3)

ggplot(s.crtsplit, aes(x=CRT.medsplit, y=QCORRECT, fill=TREATMENT)) + 
    geom_bar(stat="identity", position=dodge) + 
    geom_errorbar(limits, position=dodge, width=0.5)
```

# And looking at correlations with final exam grades
```{r}
s.fLMH <- summarySE(dat.trt, measurevar="QCORRECT",
                              groupvars=c("final.grade.LMH","TREATMENT"), .drop=FALSE)
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
#options(repr.plot.width=5, repr.plot.height=3)

ggplot(s.fLMH, aes(x=final.grade.LMH, y=QCORRECT, fill=TREATMENT)) + 
    geom_bar(stat="identity", position=dodge) + 
    geom_errorbar(limits, position=dodge, width=0.5)
```

# Finally, let's look at gender

### Bar graph splitting on gender and looking at the effect of treatment

Splitting on gender shows that there may be a significant difference in who the intervention helps
```{r}
sum.gender <- summarySE(dat.trt, measurevar="QCORRECT",
                              groupvars=c("Gender","TREATMENT"), .drop=FALSE)

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(sum.gender, aes(x=Gender, y=QCORRECT, fill=TREATMENT)) + 
    geom_bar(stat="identity", position=dodge) + 
    geom_errorbar(limits, position=dodge, width=0.5)
```


### Logistic regression with gender added

Gender is significant.

```{r}
m.gender0 <- glmer(QCORRECT ~ QNUM + TREATMENT + Gender + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.gender0)
```


### Run logistic regressions on the two gender populations seperately 

First without course grades included. Treatment is only significant for females, not males.

```{r}
m.female <- glmer(QCORRECT ~ QNUM + TREATMENT + (1|ID), 
                data = subset(dat.trt, Gender=="Female"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.female)
m<-m.female
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

```{r}
m.male <- glmer(QCORRECT ~ QNUM + TREATMENT + (1|ID), 
                data = subset(dat.trt, Gender=="Male"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.male)
m<-m.male
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

When controlling for course grade, nothing changes

```{r}
m.female.grade <- glmer(QCORRECT ~ QNUM + TREATMENT + course.grade.frac + (1|ID), 
                data = subset(dat.trt,Gender=="Female"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.female.grade)
```

```{r}
m.male.grade <- glmer(QCORRECT ~ QNUM + TREATMENT + course.grade.frac + (1|ID), 
                data = subset(dat.trt,Gender=="Male"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.male.grade)
```

### Looking at interaction term

```{r}
m.gender <- glmer(QCORRECT ~ QNUM + TREATMENT*Gender + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.gender)
m<-m.gender
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```


### Control for overall exam score (yes, re-calcuated for each question to not include the specific question in the score)

Treatment does not change significantly, but when controlling for final exam score the gender effect goes down significantly

```{r}
m.fgradeA <- glmer(QCORRECT ~ QNUM + TREATMENT + final.gradeA.fix + Gender + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.fgradeA)
```

```{r}
m.fgrade <- glmer(QCORRECT ~ QNUM + TREATMENT + final.grade.fix + Gender + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.fgrade)
```

```{r}
m.fgrade.int <- glmer(QCORRECT ~ QNUM + TREATMENT*Gender + final.grade.fix + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
summary(m.fgrade.int)
m<-m.fgrade.int
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```