---
title: "A02g Gender Correlations (P101 Final exam - Jan, 2018)"
author: "Joss Ives"
date: "June 7, 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview
This is a companion report for "A02 Initial Explain Your Answee Report (P101 Final exam - Jan, 2018)". In this report, correlations between the treatment and gender are examined in more detail. It has the same setup section as A02

# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat.csv")
```
### Add additional calculated values
```{r echo=FALSE}
dat.raw$course.grade.frac <- dat.raw$course.grade/100.
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$final.grade.LMH <- as.integer(
  cut(dat.raw$f.tot78, quantile(dat.raw$f.tot78, probs=0:3/3), include.lowest=TRUE)
  )
dat.raw$final.grade.fix <- (dat.raw$f.tot78-2.*dat.raw$QCORRECT)/76.
dat.raw$final.gradeA.fix <- (dat.raw$f.Atot40-2.*dat.raw$QCORRECT)/38.
dat.raw$f.tot100 <- dat.raw$f.tot78/.78
# Relevel
dat.raw$CONTROL <- 1-dat.raw$TREATMENT
dat.raw$GRELEVEL <- factor(dat.raw$Gender, levels=c("Male","Female"))
```
The following is a summary of the variables present in the data file. The final 4 were caculated in this notebook. "Fix" refers to removing from the overall score on the test, the score of the specific question.
```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Set categorical variables
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$QNUM      <- factor(dat.raw$QNUM)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$CONTROL <- factor(dat.raw$CONTROL)
dat.raw$f.version <- factor(dat.raw$f.version)

```

```{r echo=FALSE}
### Make some data subsets

# Keep only those data points where TREATMENT was set
dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Remove students that didn't receive full participation credit for both EYA questions
dat.all.EYA <- subset(dat.all, EYAfinal==1)

# Look only at the 4 questions that had TREATMENT
dat.trt <- subset(dat.all.EYA, QNUM==5 | QNUM==6 | QNUM==9 | QNUM==10)

```


```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```

# Treatment x Gender

In this section we compare making data subset for each Gender with using the TREATMENT x Gender interaction terms

### Run the regressions on the two groups individually

Female-only data first
```{r echo=FALSE}
m.female.data <- glmer(QCORRECT ~ TREATMENT + final.grade.fix + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, Gender=="Female"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.female.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf11 <- round(exp(tab[[2]]),3)
tf12 <- round(exp(tab[[5]]),3)
tf13 <- round(exp(tab[[8]]),3)
```

Male-only data
```{r echo=FALSE}
m.male.data <- glmer(QCORRECT ~ TREATMENT + final.grade.fix + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, Gender=="Male"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.male.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

# Grab values for reporting table
tf21 <- round(exp(tab[[2]]),3)
tf22 <- round(exp(tab[[5]]),3)
tf23 <- round(exp(tab[[8]]),3)

```

Gender effect for control condition only, controlling for question difficulty and test perfromance
```{r echo=FALSE}
m.control.data <- glmer(QCORRECT ~ Gender + final.grade.fix + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, TREATMENT==0), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.control.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf31 <- round(exp(tab[[2]]),3)
tf32 <- round(exp(tab[[5]]),3)
tf33 <- round(exp(tab[[8]]),3)
```

Gender effect for treatment condition only, controlling for question difficulty and test performance
```{r echo=FALSE}
m.control.data <- glmer(QCORRECT ~ Gender + final.grade.fix + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, TREATMENT==1), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.control.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf41 <- round(exp(tab[[2]]),3)
tf42 <- round(exp(tab[[5]]),3)
tf43 <- round(exp(tab[[8]]),3)
```

### Run the regressions on the entire data set with TREATMENT x Gender interaction term

First run with Gender=Female and TREATMENT=Control as the base levels
```{r echo=FALSE}
m.gi.f.c <- glmer(QCORRECT ~ TREATMENT*Gender + final.grade.fix + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf14 <- round(exp(tab[[2]]),3)
tf15 <- round(exp(tab[[7]]),3)
tf16 <- round(exp(tab[[12]]),3)
tf34 <- round(exp(tab[[3]]),3)
tf35 <- round(exp(tab[[8]]),3)
tf36 <- round(exp(tab[[13]]),3)
```

Then run with Gender=Male and TREATMENT=Control as the base levels
```{r echo=FALSE}


m.gi.f.c <- glmer(QCORRECT ~ TREATMENT*GRELEVEL + final.grade.fix + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
tf24 <- round(exp(tab[[2]]),3)
tf25 <- round(exp(tab[[7]]),3)
tf26 <- round(exp(tab[[12]]),3)
```


Then run with Gender=Male and TREATMENT=Treatment as the base levels
```{r echo=FALSE}


m.gi.f.c <- glmer(QCORRECT ~ CONTROL*GRELEVEL + final.grade.fix + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

And then for completion run with Gender=Female and TREATMENT=Treatment as the base levels
```{r echo=FALSE}


m.gi.f.t <- glmer(QCORRECT ~ CONTROL*Gender + final.grade.fix + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.t
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
tf44 <- round(exp(tab[[3]]),3)
tf45 <- round(exp(tab[[8]]),3)
tf46 <- round(exp(tab[[13]]),3)
```

### Summary of results

_Format table in markdown instead of R for fun_

None of the cross terms were statistically significant

| ODDS RATIOS [95%CI] | Data Subset           | Interaction Term  |
| ------------- |------------|-----|
| Treatment (Female) | `r tf11`&ast; [`r tf12`, `r tf13`] | `r tf14`&ast; [`r tf15`, `r tf16`] |
| Treatment (Male) | `r tf21` [`r tf22`, `r tf23`]  |`r tf24` [`r tf25`, `r tf26`] |
| Gender (Control) | `r tf31`&ast;&ast;&ast; [`r tf32`, `r tf33`]      | `r tf34`&ast;&ast;&ast; [`r tf35`, `r tf36`] |
| Gender (Treatment) | `r tf41`&ast;&ast;&ast; [`r tf42`, `r tf43`]      | `r tf44`&ast;&ast;&ast; [`r tf45`, `r tf46`] |



