---
title: "A04 Explain Your Answer Report (All P100/P101 Exams)"
author: "Joss Ives"
date: "June 10, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview
This report discusses the initial analysis of the W2017-T1 data from the Physics 100 course. In this course, 4 questions were used to look at the effect of asking students to explain their answer after a multiple-choice question. This used a crossover protocol, where there were 2 versions of the test and each version had 2 explain your answer questions that the other group did not. 

# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat_all.csv")
```

```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Create derived variables and set categorical ones
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$ExamGrade.fix.LMH <- as.integer(
  cut(dat.raw$ExamGrade.fix.z, 
      quantile(dat.raw$ExamGrade.fix.z, probs=0:3/3), 
      include.lowest=TRUE))
#
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$COURSE    <- factor(dat.raw$COURSE)
dat.raw$TERM      <- factor(dat.raw$TERM)
dat.raw$EXAM      <- factor(dat.raw$EXAM)
dat.raw$QNUM      <- factor(dat.raw$QNUM, 
                            levels=c("m0171.3","m0171.4","m0171.7","m0171.8",
                                     "f0171.5","f0171.6","f0171.13","f0171.14",
                                     "m1172.5","m1172.6",
                                     "f1172.5","f1172.6","f1172.9","f1172.10"
                                     ))
dat.raw$QNUM.0    <- factor(dat.raw$QNUM.0)
dat.raw$QNUM.long <- factor(dat.raw$QNUM.long)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$Version <- factor(dat.raw$Version)
dat.raw$ExamGrade.fix.LMH.factor <- dat.raw$ExamGrade.fix.LMH

```

```{r echo=FALSE}
### Make some data subsets

# Keep only those data points where TREATMENT was set
#dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Remove students that didn't receive full participation credit for both EYA questions
dat.EYA <- subset(dat.raw, EYAinclude==1)

# Remove the NAN gender folks
dat.trt <- subset(dat.EYA, Gender=="Female" | Gender=="Male")
```


```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```

```{r}
#dat.trt
```


# Data description

### How well was each question answered?

```{r echo=FALSE}
corr.by.question <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","QNUM")
  )

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
corr.by.question
```


```{r echo=FALSE}

ggplot(corr.by.question, aes(x=QNUM, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Question number", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  #theme(axis.text = element_text(face = "bold")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") + 
  guides(fill=guide_legend(title=NULL))
  
```


### Performance on the EYA questions by gender.
```{r echo=FALSE}

corr.by.gender <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","Gender")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.gender, aes(x=Gender, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Gender", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  theme(axis.text = element_text(face = "bold")) + 
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))
  
```

### Effect of intervention by exam performance group (tertile)
Since we know that females scored lower on the exams than males, perhaps there will be a difference in the effectiveness of the intervention if we look at low, medium and high performers from exams. The graph shows the largest effect on H, then L, then M

This will be investigated further in A04t ("t" for tertile)

```{r echo=FALSE}

s.fLMH <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("ExamGrade.fix.LMH.factor","TREATMENT")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(s.fLMH, aes(x=ExamGrade.fix.LMH.factor, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Final Exam Score (Tertile = Low, Medium, High)", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  #theme(legend.position="none")
  #scale_x_discrete(labels=c("1"="Low","Medium","High")) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))

```



# Quick analyses of effects sizes

### Fisher's Exact for the four study questions combined

This is a sanity check for the logistic regressions to come. Fisher's Exact Test shows an odds ration consistent with 0

```{r echo=FALSE}
## fisher.2vector
# 2 vectors of 0s and 1s are passed
# * The first is control (top row)
# * The second is treatement (bottom row)
# Adding parenthesis around assigning a variable causes the output to be displayed
(f.result <- fisher.2vector(
  dat.trt$QCORRECT[dat.trt$TREATMENT==0],
  dat.trt$QCORRECT[dat.trt$TREATMENT==1]
)) 

```

### Simple logistic regression

```{r echo=FALSE}
m.simple <- glmer(QCORRECT ~ TREATMENT + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.simple
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```


```{r echo=FALSE}
# Create df to hold results
Model.vars <- c('Simple')
AIC.val <- c(round(AIC(m),1))
Odds.treatment <- c(round(exp(tab[[2]]),3))
LL <- c(round(exp(tab[[4]]),1))
UL <- c(round(exp(tab[[6]]),1))
Significant.Vars <-('None')
r.df <- data.frame(Model.vars, AIC.val, Odds.treatment, LL, UL, Significant.Vars, stringsAsFactors = FALSE)
r.df
```



### A quick summary of odds ratios so far
```{r}
odds.table <- matrix(c(
  f.result$estimate[[1]],
  cohens.d.from.odds.simple(f.result$estimate[[1]]),
  exp(tab[[2]]),
  cohens.d.from.odds.simple(exp(tab[[2]]))
),ncol=2,byrow=TRUE)
colnames(odds.table) <- c("Odds Ratio","Cohen's d")
rownames(odds.table) <- c("Fisher's Exact Test","Logistic Regression")
as.table(odds.table)


#cat("Cohen's d\n ", cohens.d.from.odds.simple(result$estimate[[1]]),"\n")
```

# Question by question results, ordered by question easiness

```{r}
#does.this.work <- summarySE(
#  dat.trt, measurevar="ExamGrade.fix.z", groupvars=c("QNUM","TREATMENT","QCORRECT")
#  )

#limits <- aes(ymax = ExamGrade.fix.z + binomial.error, ymin = QCORRECT - binomial.error)

#does.this.work

# Summarize a dataset by two variables
# Note the use of the '.' function to allow
# group and sex to be used without quoting
#ddply(dat.trt, .(QNUM,TREATMENT,QCORRECT), summarise,
# cnt = length(QCORRECT)
#)

# An example using a formula for .variables
#ddply(baseball[1:100,], ~ year, nrow)
# Applying two functions; nrow and ncol
#ddply(baseball, .(lg), c("nrow", "ncol"))

# Calculate mean runs batted in for each year
#rbi <- ddply(baseball, .(year), summarise,
#  mean_rbi = mean(rbi, na.rm = TRUE))
# Plot a line chart of the result
#plot(mean_rbi ~ year, type = "l", data = rbi)
#
# make new variable career_year based on the
# start year for each player (id)
#base2 <- ddply(baseball, .(id), mutate,
# career_year = year - min(year) + 1
#)
```


```{r}
#qnames <- levels(dat.trt$QNUM)
#testy <- apply()
#qvec.c.0 <- length(dat.trt$QCORRECT[dat.trt$QCORRECT==0 & dat.trt$TREATMENT==0 & dat.trt$QNUM==qnames[1]])
#qvec.c.1 <- length(dat.trt$QCORRECT[dat.trt$QCORRECT==1 & dat.trt$TREATMENT==0 & dat.trt$QNUM==qnames[1]])
#qvec.t.0 <- length(dat.trt$QCORRECT[dat.trt$QCORRECT==0 & dat.trt$TREATMENT==1 & dat.trt$QNUM==qnames[1]])
#qvec.t.1 <- length(dat.trt$QCORRECT[dat.trt$QCORRECT==1 & dat.trt$TREATMENT==1 & dat.trt$QNUM==qnames[1]])
#dat.trt$QCORRECT[dat.trt$QNUM==qnames]
#dat.trt$QCORRECT[dat.trt$TREATMENT==0],
#dat.trt$QCORRECT[dat.trt$TREATMENT==1]
```



# Other non-interaction models

### Treatment + Gender

```{r echo=FALSE}
m.gender <- glmer(QCORRECT ~ TREATMENT + Gender + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.gender
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Gender', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'All'))
#r.df
```

### TREATMENT + Final Exam Grade (corrected for intervention question scores)

```{r echo=FALSE}
m.final.exam.grade <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.final.exam.grade
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Final Exam Grade', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'All'))
```

### TREATMENT + Final Exam Grade Tertile

```{r echo=FALSE}
m.final.exam.LMH <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.LMH + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.final.exam.LMH
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Final Exam LMH', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'All'))
```

### TREATMENT + Gender + Final Exam Grade (corrected for intervention question scores)

```{r echo=FALSE}
m.gender.final.exam <- glmer(QCORRECT ~ TREATMENT + Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.gender.final.exam
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Gender + Exam Grade', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[6]]),3), 
                     round(exp(tab[[10]]),3),'All'))
```

### Results

With the final exam data, we see no significant results from treatment

```{r}
r.df
```
