---
title: "A04c CRT correlations (P101 Exams)"
author: "Joss Ives"
date: "June 10, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview



# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
# For plot_model
library(sjPlot)
#
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat_all.csv")
```

```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Create derived variables and set categorical ones
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$CRT.medsplit <- factor(dat.raw$CRT.medsplit)
dat.raw$CRT.medsplit.relevel <- factor(dat.raw$CRT.medsplit,c("1","0"))
dat.raw$ExamGrade.fix.LMH <- as.integer(
  cut(dat.raw$ExamGrade.fix.z, 
      quantile(dat.raw$ExamGrade.fix.z, probs=0:3/3), 
      include.lowest=TRUE))
#
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$COURSE    <- factor(dat.raw$COURSE)
dat.raw$TERM      <- factor(dat.raw$TERM)
dat.raw$EXAM      <- factor(dat.raw$EXAM)
dat.raw$QNUM.relevel <- factor(dat.raw$QNUM, 
                            levels=c("m0171.3","m0171.4","m0171.7","m0171.8",
                                     "f0171.5","f0171.6","f0171.13","f0171.14",
                                     "m1172.5","m1172.6",
                                     "f1172.5","f1172.6","f1172.9","f1172.10"
                                     ))
dat.raw$QNUM <- factor(dat.raw$QNUM) 
dat.raw$QNUM.0    <- factor(dat.raw$QNUM.0)
dat.raw$QNUM.long <- factor(dat.raw$QNUM.long)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$Version <- factor(dat.raw$Version)
dat.raw$NCRT.factor <- factor(dat.raw$NCRT)

### Make some data subsets

# Remove students that didn't receive full participation credit for both EYA questions
dat.EYA <- subset(dat.raw, EYAinclude==1)
# Include only P101 data, and only if those data have CRT data
#dat.p101 <- subset (dat.EYA, COURSE=="P101")
#dat.CRT <- subset(dat.p101, NCRT==0 | NCRT==1 | NCRT==2 | NCRT==3)
dat.CRT <- subset(dat.EYA, NCRT==0 | NCRT==1 | NCRT==2 | NCRT==3)
# Remove the NAN gender folks
dat.trt <- dat.CRT
dat.trt.gender <- subset(dat.trt, Gender=="Female" | Gender=="Male")
```


```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```


# Data description

### How well was each question answered?

```{r echo=FALSE}
corr.by.question <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","QNUM.relevel")
  )

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
corr.by.question
```


```{r echo=FALSE}

ggplot(corr.by.question, aes(x=QNUM.relevel, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Question number", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  #theme(axis.text = element_text(face = "bold")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") + 
  guides(fill=guide_legend(title=NULL))
  
```


### Performance on the EYA questions by gender.

This comparison is not really that important considering what we ultimately figure out about the lack of gender effects, but we leave it for now.

```{r echo=FALSE}

corr.by.gender <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","Gender")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.gender, aes(x=Gender, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Gender", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  theme(axis.text = element_text(face = "bold")) + 
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))
  
```

### Effect of intervention by exam performance (tertile)

As compared to the overall results from A04, we see that for the P101 exams, the intervention favours L more than H more than M (overall is H > L > M)

```{r echo=FALSE}

s.fLMH <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("ExamGrade.fix.LMH","TREATMENT")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(s.fLMH, aes(x=ExamGrade.fix.LMH, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Final Exam Score (Tertile = Low, Medium, High)", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  #theme(legend.position="none")
  #scale_x_discrete(labels=c("Low","Medium","High")) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))

```

### Effect of intervention by CRT score and CRT median split

```{r echo=FALSE}

corr.by.CRT <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","NCRT")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.CRT, aes(x=NCRT, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "CRT Score", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  theme(axis.text = element_text(face = "bold")) + 
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))
  
```

```{r echo=FALSE}

corr.by.CRTmedsplit <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","CRT.medsplit")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.CRTmedsplit, aes(x=CRT.medsplit, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "CRT Score Median Split", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  theme(axis.text = element_text(face = "bold")) + 
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))
  
```

# Logsitic regressions

### Run on both of the median splits

To help make sure that the interaction terms are being interpreted correctly.

CRT Low


```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, CRT.medsplit==0), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

CRT High

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, CRT.medsplit==1), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

### Run with interaction terms

CRT low as baseline

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT*CRT.medsplit + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt,
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

CRT high as baseline

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT*CRT.medsplit.relevel + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt,
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```



### Try to grab useful variables from the fit

```{r}

length(fixef(m))
names(fixef(m))

fixef.df <-as.data.frame(fixef(m))
#fixef.df
rownames(fixef.df)
fixef.df["TREATMENT1",]
exp(tab)
odds.df <-as.data.frame(exp(tab))
odds.df["TREATMENT1","Est"]
odds.df["TREATMENT1","LL"]
odds.df["TREATMENT1","UL"]
```

```{r}
coef(summary(m))
```

```{r}
plot_model(m, vline.color = "blue")
```

