---
title: "A04g Gender Interactions (All P100/P101 Exams)"
author: "Joss Ives"
date: "June 10, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview
This report discusses the initial analysis of the W2017-T1 data from the Physics 100 course. In this course, 4 questions were used to look at the effect of asking students to explain their answer after a multiple-choice question. This used a crossover protocol, where there were 2 versions of the test and each version had 2 explain your answer questions that the other group did not. 

# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat_all.csv")
```

```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Set categorical variables
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$COURSE    <- factor(dat.raw$COURSE)
dat.raw$TERM      <- factor(dat.raw$TERM)
dat.raw$EXAM      <- factor(dat.raw$EXAM)
dat.raw$QNUM      <- factor(dat.raw$QNUM)
dat.raw$QNUM.0    <- factor(dat.raw$QNUM.0)
dat.raw$QNUM.long <- factor(dat.raw$QNUM.long)
dat.raw$CONTROL   <- 1-dat.raw$TREATMENT
dat.raw$CONTROL   <- factor(dat.raw$CONTROL)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$Version   <- factor(dat.raw$Version)
dat.raw$GRELEVEL  <- factor(dat.raw$Gender, levels=c("Male","Female"))
```

```{r echo=FALSE}
### Make some data subsets

# Keep only those data points where TREATMENT was set
#dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Remove students that didn't receive full participation credit for both EYA questions
dat.EYA <- subset(dat.raw, EYAinclude==1)

# Remove the NAN gender folks
dat.trt <- subset(dat.EYA, Gender=="Female" | Gender=="Male")
```


```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```



### Run the regressions on the two groups individually

Female-only data first
```{r echo=FALSE}
m.female.data <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, Gender=="Female"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.female.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf11 <- round(exp(tab[[2]]),3)
tf12 <- round(exp(tab[[5]]),3)
tf13 <- round(exp(tab[[8]]),3)
```

Male-only data
```{r echo=FALSE}
m.male.data <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, Gender=="Male"), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.male.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

# Grab values for reporting table
tf21 <- round(exp(tab[[2]]),3)
tf22 <- round(exp(tab[[5]]),3)
tf23 <- round(exp(tab[[8]]),3)

```

Gender effect for control condition only, controlling for question difficulty and test perfromance
```{r echo=FALSE}
m.control.data <- glmer(QCORRECT ~ Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, TREATMENT==0), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.control.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf31 <- round(exp(tab[[2]]),3)
tf32 <- round(exp(tab[[5]]),3)
tf33 <- round(exp(tab[[8]]),3)
```

Gender effect for treatment condition only, controlling for question difficulty and test performance
```{r echo=FALSE}
m.control.data <- glmer(QCORRECT ~ Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, TREATMENT==1), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.control.data
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf41 <- round(exp(tab[[2]]),3)
tf42 <- round(exp(tab[[5]]),3)
tf43 <- round(exp(tab[[8]]),3)
```

### Run the regressions on the entire data set with TREATMENT x Gender interaction term

First run with Gender=Female and TREATMENT=Control as the base levels
```{r echo=FALSE}
m.gi.f.c <- glmer(QCORRECT ~ TREATMENT*Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

tf14 <- round(exp(tab[[2]]),3)
tf15 <- round(exp(tab[[7]]),3)
tf16 <- round(exp(tab[[12]]),3)
tf34 <- round(exp(tab[[3]]),3)
tf35 <- round(exp(tab[[8]]),3)
tf36 <- round(exp(tab[[13]]),3)
```

Then run with Gender=Male and TREATMENT=Control as the base levels
```{r echo=FALSE}


m.gi.f.c <- glmer(QCORRECT ~ TREATMENT*GRELEVEL + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
tf24 <- round(exp(tab[[2]]),3)
tf25 <- round(exp(tab[[7]]),3)
tf26 <- round(exp(tab[[12]]),3)
```

**This one is not working and it seems to be due to my additional variable CONTROL**

Then run with Gender=Male and TREATMENT=Treatment as the base levels
```{r echo=FALSE}


m.gi.f.c <- glmer(QCORRECT ~ CONTROL*GRELEVEL + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.c
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
```

And then for completion run with Gender=Female and TREATMENT=Treatment as the base levels
```{r echo=FALSE}


m.gi.f.t <- glmer(QCORRECT ~ CONTROL*Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m<-m.gi.f.t
print(summary(m))
#
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
tf44 <- round(exp(tab[[3]]),3)
tf45 <- round(exp(tab[[8]]),3)
tf46 <- round(exp(tab[[13]]),3)
```

### Summary of results

_Format table in markdown instead of R for fun_

None of the cross terms were statistically significant

| ODDS RATIOS [95%CI] | Data Subset           | Interaction Term  |
| ------------- |------------|-----|
| Treatment (Female) | `r tf11`&ast;&ast;&ast; [`r tf12`, `r tf13`] | `r tf14`&ast;&ast;&ast; [`r tf15`, `r tf16`] |
| Treatment (Male) | `r tf21`&ast; [`r tf22`, `r tf23`]  |`r tf24`&ast; [`r tf25`, `r tf26`] |
| Gender (Control) | `r tf31`&ast;&ast;&ast; [`r tf32`, `r tf33`]      | `r tf34`&ast;&ast;&ast; [`r tf35`, `r tf36`] |
| Gender (Treatment) | `r tf41`&ast;&ast;&ast; [`r tf42`, `r tf43`]      | `r tf44`&ast;&ast;&ast; [`r tf45`, `r tf46`] |

```{r}
cohens.d.from.odds.simple(tf14)
cohens.d.from.odds.simple(tf24)
```



