---
title: "A04 Explain Your Answer Report (All P100/P101 Exams)"
author: "Joss Ives"
date: "June 10, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview
This report discusses the initial analysis of the W2017-T1 data from the Physics 100 course. In this course, 4 questions were used to look at the effect of asking students to explain their answer after a multiple-choice question. This used a crossover protocol, where there were 2 versions of the test and each version had 2 explain your answer questions that the other group did not. 

# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat_all.csv")
```

```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Create derived variables and set categorical ones
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$ExamGrade.fix.LMH <- as.integer(
  cut(dat.raw$ExamGrade.fix.z, 
      quantile(dat.raw$ExamGrade.fix.z, probs=0:3/3), 
      include.lowest=TRUE))
#
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$COURSE    <- factor(dat.raw$COURSE)
dat.raw$TERM      <- factor(dat.raw$TERM)
dat.raw$EXAM      <- factor(dat.raw$EXAM)
dat.raw$QNUM      <- factor(dat.raw$QNUM, 
                            levels=c("m0171.3","m0171.4","m0171.7","m0171.8",
                                     "f0171.5","f0171.6","f0171.13","f0171.14",
                                     "m1172.5","m1172.6",
                                     "f1172.5","f1172.6","f1172.9","f1172.10"
                                     ))
dat.raw$QNUM.m0171.4 <- relevel(dat.raw$QNUM,("m0171.4"))

dat.raw$QNUM.0    <- factor(dat.raw$QNUM.0)
dat.raw$QNUM.long <- factor(dat.raw$QNUM.long)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$Version <- factor(dat.raw$Version)
dat.raw$ExamGrade.fix.LMH.factor <- dat.raw$ExamGrade.fix.LMH

```

```{r}
#levels(dat.raw$QNUM.m0171.4)
```


```{r echo=FALSE}
### Make some data subsets

# Keep only those data points where TREATMENT was set
#dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Remove students that didn't receive full participation credit for both EYA questions
dat.EYA <- subset(dat.raw, EYAinclude==1)

# Remove the NAN gender folks
dat.trt <- subset(dat.EYA, Gender=="Female" | Gender=="Male")
```


```{r}
varlist <- levels(dat.trt$QNUM)
vlen <- length(varlist)

varlist2 <- character(vlen)
odds <- numeric(vlen)
LL <- numeric(vlen)
UL <- numeric(vlen)
pval <- numeric(vlen)

for (i in 1:vlen){
  dat.trt$QNUM.r <- relevel(dat.trt$QNUM,varlist[i])
  print(i)
  dat.trt$QNUM.r
  m <- glmer(QCORRECT ~ TREATMENT*QNUM.r + ExamGrade.fix.z + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
  summary(m)
  se <- sqrt(diag(vcov(m)))
  tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
  odds.df <-as.data.frame(exp(tab))
  varlist2[i] <- varlist[i] # Sanity check
  odds[i] <-odds.df["TREATMENT1","Est"]
  LL[i] <- odds.df["TREATMENT1","LL"]
  UL[i] <- odds.df["TREATMENT1","UL"]
  fe.df <- as.data.frame(coef(summary(m)))
  pval[i] <- fe.df["TREATMENT1","Pr(>|z|)"]
  }
varlist2
odds
LL
UL
pval
```

```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```


```{r}
q <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("QNUM")
  )

obq.df <- data.frame(order.0 = as.factor(1:vlen),
                 y = odds,
                 ymin = LL,
                 ymax = UL,
                 ease = q$QCORRECT,
                 qname = varlist)

ease.df <- obq.df[order(obq.df$ease),]
ease.df$ease.order <- as.factor(1:vlen)
ease.df

obq.df2 <- ease.df[order(ease.df$order.0),]
obq.df2
```


```{r echo=FALSE}
corr.by.question <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","QNUM")
  )

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(corr.by.question, aes(x=QNUM, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Question number", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  #theme(axis.text = element_text(face = "bold")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  scale_x_discrete(limits=ease.df$qname) +
  ggtitle("Error bars are binomial errors") + 
  guides(fill=guide_legend(title=NULL))
  
obq.df <- data.frame(x = as.factor(1:vlen),
                 y = odds,
                 ymin = LL,
                 ymax = UL)

ggplot(data = ease.df,aes(x = ease.order,y = y)) + 
    geom_point() + 
    geom_errorbar(aes(ymin = ymin,ymax = ymax)) +
    geom_hline(yintercept=1, color="blue") +
    #scale_x_discrete(limits=ease.df$qname) +
    theme_bw() 
    

```

Will likely want to mix multiple plots http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/

