---
title: "A04c CRT correlations (P101 Exams)"
author: "Joss Ives"
date: "June 10, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*updated by Joss Ives `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*

# Overview



# Setup

```{r echo=FALSE}
# Load libraries
library(plyr)
library(lme4)
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
source("jossfunc.R")
```

```{r echo=FALSE}
# Load data
dat.raw <- read.csv("C:/Users/Joss/ownCloud/Shared/DP_Derived_Data/dpLogisticDat_all.csv")
```

```{r}
names(dat.raw)
```

```{r echo = FALSE}
### Create derived variables and set categorical ones
dat.raw$CRT.medsplit <- trunc(dat.raw$NCRT/2)
dat.raw$ExamGrade.fix.LMH <- as.integer(
  cut(dat.raw$ExamGrade.fix.z, 
      quantile(dat.raw$ExamGrade.fix.z, probs=0:3/3), 
      include.lowest=TRUE))
#
dat.raw$ID        <- factor(dat.raw$ID)
dat.raw$COURSE    <- factor(dat.raw$COURSE)
dat.raw$TERM      <- factor(dat.raw$TERM)
dat.raw$EXAM      <- factor(dat.raw$EXAM)
dat.raw$QNUM.relevel <- factor(dat.raw$QNUM, 
                            levels=c("m0171.3","m0171.4","m0171.7","m0171.8",
                                     "f0171.5","f0171.6","f0171.13","f0171.14",
                                     "m1172.5","m1172.6",
                                     "f1172.5","f1172.6","f1172.9","f1172.10"
                                     ))
dat.raw$QNUM <- factor(dat.raw$QNUM) 
dat.raw$QNUM.0    <- factor(dat.raw$QNUM.0)
dat.raw$QNUM.long <- factor(dat.raw$QNUM.long)
dat.raw$TREATMENT <- factor(dat.raw$TREATMENT)
dat.raw$Version <- factor(dat.raw$Version)
```

```{r echo=FALSE}
### Make some data subsets

# Keep only those data points where TREATMENT was set
#dat.all <- subset(dat.raw, TREATMENT==0 | TREATMENT==1)

# Remove students that didn't receive full participation credit for both EYA questions
dat.EYA <- subset(dat.raw, EYAinclude==1)
# Include only P101 data, and only if those data have CRT data
dat.trt <- subset(dat.EYA, Gender=="Female" | Gender=="Male")
```


```{r echo = FALSE}
### Set some global variables
dodge=position_dodge(width=0.9)
```

```{r}
#dat.trt
```


# Data description

### How well was each question answered?

```{r echo=FALSE}
corr.by.question <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("TREATMENT","QNUM.relevel")
  )

limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)
corr.by.question
```


```{r echo=FALSE}

ggplot(corr.by.question, aes(x=QNUM.relevel, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Question number", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  theme_bw() + 
  #theme(axis.text = element_text(face = "bold")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #theme(legend.position="none")
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") + 
  guides(fill=guide_legend(title=NULL))
  
```


### Effect of intervention by exam performance (tertile)

```{r echo=FALSE}

s.fLMH <- summarySE(
  dat.trt, measurevar="QCORRECT", groupvars=c("ExamGrade.fix.LMH","TREATMENT")
  )
limits <- aes(ymax = QCORRECT + binomial.error, ymin = QCORRECT - binomial.error)

ggplot(s.fLMH, aes(x=ExamGrade.fix.LMH, y=QCORRECT, fill=TREATMENT)) +
  geom_bar(stat="identity", position=dodge) + 
  geom_errorbar(limits, position=dodge, width=0.25) +
  labs(x = "Final Exam Score (Tertile = Low, Medium, High)", y = "Fraction of EYA questions correct") +
  scale_fill_discrete(labels=c("Control","Treatment")) +
  #theme(legend.position="none")
  #scale_x_discrete(labels=c("Low","Medium","High")) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("Error bars are binomial errors") +
  guides(fill=guide_legend(title=NULL))

```

# Logsitic regressions

### Run on all three LMH data subsets 

To help make sure that the interaction terms are being interpreted correctly.

Exam performance = "L" ("1")

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, ExamGrade.fix.LMH==1), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

odds.df <-as.data.frame(exp(tab))
name1 <- "L"
val1 <-odds.df["TREATMENT1","Est"]
LL1 <- odds.df["TREATMENT1","LL"]
UL1 <- odds.df["TREATMENT1","UL"]
fe.df <- as.data.frame(coef(summary(m)))
p1 <- fe.df["TREATMENT1","Pr(>|z|)"]

```

Exam performance = "M" ("2")

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, ExamGrade.fix.LMH==2), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)

odds.df <-as.data.frame(exp(tab))
name2 <- "M"
val2 <-odds.df["TREATMENT1","Est"]
LL2 <- odds.df["TREATMENT1","LL"]
UL2 <- odds.df["TREATMENT1","UL"]
fe.df <- as.data.frame(coef(summary(m)))
p2 <- fe.df["TREATMENT1","Pr(>|z|)"]

```

Exam performance = "H" ("3")

```{r echo=FALSE}
m <- glmer(QCORRECT ~ TREATMENT + (1|QNUM) + (1|ID), 
                data = subset(dat.trt, ExamGrade.fix.LMH==3), 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
name3 <- "H"

odds.df <-as.data.frame(exp(tab))
val3 <-odds.df["TREATMENT1","Est"]
LL3 <- odds.df["TREATMENT1","LL"]
UL3 <- odds.df["TREATMENT1","UL"]
fe.df <- as.data.frame(coef(summary(m)))
p3 <- fe.df["TREATMENT1","Pr(>|z|)"]

```

# Summarize the treatment values

```{r}

Group <- c(name1,name2,name3)
Treatment.val <-c(val1,val2,val3)
LL <-c(LL1,LL2,LL3)
UL <-c(UL1,UL2,UL3)
p.val <-c(p1,p2,p3)
data.frame(Group,Treatment.val,LL,UL,p.val)
```



###


```{r echo=FALSE}
# Create df to hold results
Model.vars <- c('Simple')
AIC.val <- c(round(AIC(m),1))
Odds.treatment <- c(round(exp(tab[[2]]),3))
LL <- c(round(exp(tab[[4]]),1))
UL <- c(round(exp(tab[[6]]),1))
Significant.Vars <-('None')
r.df <- data.frame(Model.vars, AIC.val, Odds.treatment, LL, UL, Significant.Vars, stringsAsFactors = FALSE)
r.df
```


# Other non-interaction models

### Treatment + Gender

```{r echo=FALSE}
m.gender <- glmer(QCORRECT ~ TREATMENT + Gender + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.gender
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Gender', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'***Gender'))
#r.df
```

### TREATMENT + Final Exam Grade (corrected for intervention question scores)

```{r echo=FALSE}
m.final.exam.grade <- glmer(QCORRECT ~ TREATMENT + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.final.exam.grade
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Final Exam Grade', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'None'))
```

### TREATMENT + Final Exam Grade Tertile

```{r echo=FALSE}
m.final.exam.LMH <- glmer(QCORRECT ~ TREATMENT + ExamGrade.LMH + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.final.exam.LMH
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Final Exam LMH', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[5]]),3), 
                     round(exp(tab[[8]]),3),'None'))
```

### TREATMENT + Gender + Final Exam Grade (corrected for intervention question scores)

```{r echo=FALSE}
m.gender.final.exam <- glmer(QCORRECT ~ TREATMENT + Gender + ExamGrade.fix.z + (1|QNUM) + (1|ID), 
                data = dat.trt, 
                family = binomial, control=glmerControl(optimizer="bobyqa"))
m <- m.gender.final.exam
print(summary(m))
se <- sqrt(diag(vcov(m)))
tab <- cbind(Est = fixef(m), LL = fixef(m) - 1.96 * se, UL = fixef(m) + 1.96 *se)
exp(tab)
r.df <- rbind(r.df,c('Gender + Exam Grade', round(AIC(m),1), round(exp(tab[[2]]),3), round(exp(tab[[6]]),3), 
                     round(exp(tab[[10]]),3),'None'))
```

### Results

With the final exam data, we see no significant results from treatment

```{r}
r.df
```
