---
title: "Dual processing data preparation"
output:
  html_notebook: default
---
Produced by Joss Ives, *updated `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*. Based on data preparation documents by Patrick Dubois: [GitHub](https://github.com/pajodu/GERG/)

## Overview

This file combines midterm, final exam and diagnostic data into a single file. This data format may need to be further modified for the logistic analysis.

## Load libraries
```{r}
#install.packages("dplyr")
library("dplyr")
```


## Set directories

Start by setting directories

```{r}
startDir <- getwd()
rawDir <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Raw_Data/P100 - 2017-09"
anonDir <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Anonymized_Data/P100 - 2017-09"
```

Load raw data

```{r}
dataSources <- data.frame(Filename=c(
                                      "P100-Sept_2017-Final_AnonID.csv", # Gender info
                                      "P100-final-gradesheet-EYA-mod.csv" # Final exam results
                                    ),
                            Tag=c("Classlist", "Final") )

datRawClass <- read.csv(file.path(rawDir,dataSources$Filename[1]))
datRawFinal <- read.csv(file.path(rawDir,dataSources$Filename[2]))
```

## Make a new classlist data frame with only the desired column

```{r}
names(datRawClass)
```

```{r}
curDat <- select(datRawClass, ID, AnonID, Gender)
datClass <- curDat
names(datClass)
```

## Make a new final exam data frame with only the desired columns

```{r}
names(datRawFinal)
```


```{r}
curDat <- select(datRawFinal, ID, 
                 A5.f0, A6.f0, A13.f0, A14.f0, 
                 SoloTotal74.f0, ExamGrade100.f0,
                 A5.trt.f0, A6.trt.f0, A13.trt.f0, A14.trt.f0,
                 EYAinclude.f0, version.f0)
datFinal <- curDat
names(datFinal)
```

## Merge the data sets by ID and then make an anonymous version

```{r}
dpDat <- merge(datClass,datFinal,by="ID")
dpDat <- select(dpDat,-ID)
names(dpDat)
```


Temporary cheat to see workflow. Let's just save this anonymized data set to the derived data folder

```{r}

dpOut <- "dpDat.csv"
write.csv(dpDat, file.path(anonDir,dpOut), row.names = FALSE)
```

