---
title: "Dual processing data preparation"
output:
  html_notebook: default
---
Produced by Joss Ives, *updated `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*. Based on data preparation documents by Patrick Dubois: [GitHub](https://github.com/pajodu/GERG/)

## Overview

This file combines midterm, final exam and diagnostic data into a single file. This data format may need to be further modified for the logistic analysis.

## To action

Start by setting directories

```{r}
StartDir <- getwd()
SourceDir <- "W:/Shared_Read_Only/DP_Anonymized_Data/P101 - 2018-01"
DerivedDir <- "W:/Shared/DP_Derived_Data"
```

Load anonymized data

```{r}
### This wasn't working with one file
#DataSources <- data.frame(Filename=c("Diagnostic - Raw - Anon Random.csv"), Tag=c("P101-2018-01"))
#dat1801 <- read.csv(DataSources$Filename[1])
dat1801 <- read.csv(file.path(SourceDir,"Diagnostic - Raw - Anon Random.csv"))

```



```{r}
Names1801 <-names(dat1801)
Names1801
NewNames <- ("dunno")
NameRef1801 <- data.frame(NewNames,Names1801)
```

Temporary cheat to see workflow. Let's just save this anonymized data set to the derived data folder

```{r}
write.csv(dat1801, file.path(DerivedDir,"test.csv"), row.names = FALSE)
```

