---
title: "Dual processing data preparation"
output:
  html_notebook: default
---
Produced by Joss Ives, *updated `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*. Based on data preparation documents by Patrick Dubois: [GitHub](https://github.com/pajodu/GERG/)

## Overview

This file combines midterm, final exam and diagnostic data into a single file. This data format may need to be further modified for the logistic analysis.

## Load libraries
```{r}
#install.packages("dplyr")
#library("dplyr")
```


## Set directories

Start by setting directories

```{r}
startDir <- getwd()
#SourceDir <- "W:/Shared_Read_Only/DP_Anonymized_Data/P101 - 2018-01"
rawDir <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Raw_Data/P101 - 2018-01"
#DerivedDir <- "W:/Shared/DP_Derived_Data"
#DerivedDir <- "C:/Users/Joss/ownCloud/Shared/DP_Derived_Data"
anonDir <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Anonymized_Data/P101 - 2018-01"
```

Load raw data

```{r}
### This wasn't working with one file
#DataSources <- data.frame(Filename=c("Diagnostic - Raw - Anon Random.csv"), Tag=c("P101-2018-01"))
#dat1801 <- read.csv(DataSources$Filename[1])
dataSources <- data.frame(Filename=c(
                                      "Classlist.csv", 
                                      "Diagnostic - Raw - Not Anon.csv",
                                      "MT solo-gradesheet-graded_tab.csv",
                                      "solo-gradesheet (final) graded_tab.csv"
                                    ),
                            Tag=c("Classlist", "Diagnostic", "Midterm","Final") )


#DatClass <- read.csv(paste0(RawDir,"/",DataSources$Filename[1]))
datRawClass <- read.csv(file.path(rawDir,dataSources$Filename[1]))
datRawDiagn <- read.csv(file.path(rawDir,dataSources$Filename[2]))
datRawMT <- read.csv(file.path(rawDir,dataSources$Filename[3]))
datRawFinal <- read.csv(file.path(rawDir,dataSources$Filename[4]))
```

## Make a new classlist data frame with only the desired column

```{r}
names(datRawClass)
```

```{r}
curDat <- select(datRawClass, ID, Gender,Percent.Grade)
# Rename a Percent.Grade
colnames(curDat)[colnames(curDat)=="Percent.Grade"] <- "Course.Grade"
#namesClass <-names(datClass)
#namesClass
#NameRef1801 <- data.frame(NewNames,Names1801)
datClass <- curDat
names(datClass)
```

## Make a new diagnostic data frame with only the desired columns

```{r}
names(datRawDiagn)
```


```{r}
curDat <- select(datRawDiagn, ID, AnonID, Version, NCRT
                 )
colnames(curDat)[colnames(curDat)=="Version"] <- "d.version"
datDiagn <- curDat
names(datDiagn)
```


## Make a new final exam data frame with only the desired columns

```{r}
names(datRawFinal)
```


```{r}
curDat <- select(datRawFinal, ID, FinalVersion, 
                 f.A1, f.A2, f.A3, f.A4, f.A5, 
                 f.A6, f.A7, f.A8, f.A9, f.A10, 
                 f.A11,f.A12,f.A13,f.A14,f.A15,
                 f.A16,f.A17,f.A18,f.A19a,f.A19b, f.Atotal40,
                 f.B1, f.B2, f.B3, f.B4, f.Btotal38, f.Solo78,
                 f.A5.trt, f.A6.trt, f.A9.trt, f.A10.trt)
colnames(curDat)[colnames(curDat)=="FinalVersion"] <- "f.version"
datFinal <- curDat
names(datFinal)
```

## Merge the data sets by ID and then make an anonymous version

```{r}
dpDat <- merge(datClass,datDiagn,by="ID")
dpDat <- merge(dpDat,datFinal,by="ID")
dpDat <- select(dpDat,-ID)
names(dpDat)
```


Temporary cheat to see workflow. Let's just save this anonymized data set to the derived data folder

```{r}

dpOut <- "dpDat.csv"
write.csv(dpDat, file.path(anonDir,dpOut), row.names = FALSE)
```

