---
title: "Dual processing data preparation"
output:
  html_notebook: default
---
Produced by Joss Ives, *updated `r format(Sys.time(), "%Y %B %d, %H:%M:%S")`*. Based on data preparation documents by Patrick Dubois: [GitHub](https://github.com/pajodu/GERG/)

## Overview

This file combines midterm, final exam and diagnostic data into a single file. This data format may need to be further modified for the logistic analysis.

## Load libraries
```{r}
#install.packages("dplyr")
library("dplyr")
```


## Set directories

Start by setting directories

```{r}
startDir <- getwd()
rawDir100 <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Raw_Data/P100 - 2017-09"
anonDir100 <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Anonymized_Data/P100 - 2017-09"
rawDir101 <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Raw_Data/P101 - 2018-01"
anonDir101 <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Anonymized_Data/P101 - 2018-01"
genderDir <- "C:/Users/Joss/ownCloud/Shared_Read_Only/DP_Raw_Data"
```

Load raw data

```{r}
dataSources <- data.frame(Filename=c(
                                      "P10X-Anon_and_Gender.csv", #Gender and anon ID info
                                      "P100-midterm-gradesheet-EYA.csv", # P100 Midterm
                                      "P100-final-gradesheet-EYA-mod.csv", # P100 Final exam results
                                      "Diagnostic - Raw - Not Anon.csv", # P101 diagnostic
                                      "MT solo-gradesheet-graded_tab-EYA.csv", #P101 Midterm
                                      "solo-gradesheet (final) graded_tab.csv" # P101 Final
                                    ),
                            Tag=c("Classlist", "P100MT", "P100final", "P101diagnostic", "P101MT", "P101final") )

datRawClass <- read.csv(file.path(genderDir,dataSources$Filename[1]))
datRawMT100 <- read.csv(file.path(rawDir100,dataSources$Filename[2]))
datRawFinal100 <- read.csv(file.path(rawDir100,dataSources$Filename[3]))
datRawDiagnostic <- read.csv(file.path(rawDir101,dataSources$Filename[4]))
datRawMT101 <- read.csv(file.path(rawDir101,dataSources$Filename[5]))
datRawFinal101 <- read.csv(file.path(rawDir101,dataSources$Filename[6]))
```

## Make a classlist df with only the desired columns

```{r}
names(datRawClass)
```

```{r}
curDatClass <- select(datRawClass, ID, AnonID, Gender) 
names(curDatClass)
```

## Make a MT100 df with only the desired columns

```{r}
names(datRawMT100)
```

```{r}
curDatMT100 <- select(datRawMT100, ID, 
                      A3.score, A4.score, A7.score, A8.score,
                      A3.trt, A4.trt, A7.trt, A8.trt,
                      SoloTotal36, EYAinclude)
# Rename question numbers
colnames(curDatMT100)[colnames(curDatMT100)=="A3.score"] <- "A3"
colnames(curDatMT100)[colnames(curDatMT100)=="A4.score"] <- "A4"
colnames(curDatMT100)[colnames(curDatMT100)=="A7.score"] <- "A7"
colnames(curDatMT100)[colnames(curDatMT100)=="A8.score"] <- "A8"

names(curDatMT100)
```

### Merge the data sets by ID and then make an anonymous version

```{r}
dpDatP100midterm <- merge(curDatMT100,curDatClass,by="ID")
dpDatP100midterm <- select(dpDatP100midterm,-ID)
names(dpDatP100midterm)
```

### Save file to anonymized folder

```{r}

dpOut <- "dpDatP100midterm.csv"
write.csv(dpDatP100midterm, file.path(anonDir100,dpOut), row.names = FALSE)
```


## Make a Final100 df with only the desired columns

```{r}
names(datRawFinal100)
```

```{r}
curDatFinal100 <- select(datRawFinal100, ID, 
                      A5, A6, A13, A14,
                      A5.trt, A6.trt, A13.trt, A14.trt,
                      SoloTotal74, EYAinclude)
names(curDatFinal100)
```

### Merge the data sets by ID and then make an anonymous version

```{r}
dpDatP100final <- merge(curDatFinal100,curDatClass,by="ID")
dpDatP100final <- select(dpDatP100final,-ID)
names(dpDatP100final)
```

### Save file to anonymized folder

```{r}

dpOut <- "dpDatP100final.csv"
write.csv(dpDatP100final, file.path(anonDir100,dpOut), row.names = FALSE)
```


## Make a MT101 df with only the desired columns

```{r}
names(datRawMT101)
```

```{r}
curDatMT101 <- select(datRawMT101, ID, 
                      A5.score, A6.score,
                      A5.trt, A6.trt,
                      SoloTotal43, EYAinclude)
# Rename question numbers
colnames(curDatMT101)[colnames(curDatMT101)=="A5.score"] <- "A5"
colnames(curDatMT101)[colnames(curDatMT101)=="A6.score"] <- "A6"
names(curDatMT101)
```

### Merge the data sets by ID and then make an anonymous version

```{r}
dpDatP101midterm <- merge(curDatMT101,curDatClass,by="ID")
dpDatP101midterm <- select(dpDatP101midterm,-ID)
names(dpDatP101midterm)
```

### Save file to anonymized folder

```{r}

dpOut <- "dpDatP101midterm.csv"
write.csv(dpDatP101midterm, file.path(anonDir101,dpOut), row.names = FALSE)
```


## Make a Final101 df with only the desired columns

```{r}
names(datRawFinal101)
```

```{r}
curDatFinal101 <- select(datRawFinal101, ID, 
                      A5, A6, A9, A10,
                      A5.trt, A6.trt, A9.trt, A10.trt,
                      SoloTotal78, EYAinclude)
names(curDatFinal101)
```

### Merge the data sets by ID and then make an anonymous version

```{r}
dpDatP101final <- merge(curDatFinal101,curDatClass,by="ID")
dpDatP101final <- select(dpDatP101final,-ID)
names(dpDatP101final)
```

### Save file to anonymized folder

```{r}

dpOut <- "dpDatP101final.csv"
write.csv(dpDatP101final, file.path(anonDir101,dpOut), row.names = FALSE)
```








